---
title: "Zimbabwe Soybean Trials"
author: "José Tiago"
date: "`r Sys.Date()`"
always_allow_html: yes
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading data
```{r}
#install.packages("sf")
#install.packages("stars")
library(tidyverse) 
data <- read.csv("data/Zimbabwe.csv",header = T,sep = ";")
colnames(data)
head(data)
pheno<- data[, c(4, 7, 16,21,25,27,38,41,46,63,64,65,66,68)] 
table(pheno$LOCATION_NAME)
colnames(pheno)
pheno$geno <- gsub(" ", "_",pheno$Crop_Variety)
pheno$geno <- factor(pheno$geno)
unique(pheno$IRRIGATION)
levels(pheno$geno) <- paste("G", 1:length(levels(pheno$geno)), sep = "_")
table(pheno$Crop_Variety)
table(pheno$geno)

pheno[which(pheno$HARV_DATE==""),"LOCATION_NAME"]

pheno$DTH <- as.Date(as.character(pheno$HARV_DATE), format="%m/%d/%Y")-
                  as.Date(as.character(pheno$PLANT_DATE), format="%m/%d/%Y")-5
dim(pheno)
colnames(pheno)
head(pheno)
str(pheno)
pheno$GRAIN_YLD_13PCT <-  gsub(",", ".", pheno$GRAIN_YLD_13PCT)
trial <-  gsub("_.*", "", pheno$LOCATION_NAME)
pheno$trial <- paste0(trial,"_", pheno$CROP_SEASON)
unique(pheno$TRIAL_NAME) # 41 trials
unique(pheno$LOCATION_NAME) # 14 Locations
unique(pheno$CROP_SEASON) # 6 years
length(unique(pheno$Crop_Variety)) #98 genotypes
head(pheno)
pheno <- pheno |> rename("Plant Height"=8, "Grain Yield"=9, "Plant lodging"=7, "rep"=4)

traits <- c(colnames(pheno)[c(7,8,9)])
traits
pheno[, traits] <- lapply(pheno[, traits], as.numeric)

factors <- c(colnames(pheno)[c(1,2,3,4,5,6,14,15,16,17)])
factors


traits <- c(colnames(pheno)[c(7,8,9)])
traits

pheno[, factors] <- lapply(pheno[, factors], as.factor)
head(pheno)

#Irrigation system
table(pheno$trial,pheno$IRRIGATION, useNA = "ifany")
unique(pheno$IRRIGATION)
pheno[which(pheno$trial=="Banket_2019"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Banket_2020"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Bindura_2020"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Chisumbanje_2019"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Goromonzi_2019"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Goromonzi_2020"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Gwebi_2020"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Mutare_2019"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial==" Mutare_2020"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Stapleford_2019"),"IRRIGATION"] <- "Rainfed"
pheno[which(pheno$trial=="Stapleford_2020"),"IRRIGATION"] <- "Rainfed"

table(pheno$IRRIGATION)


levels(pheno$IRRIGATION)
```



# Data visualization

```{r}

pheno_long <- reshape2::melt(pheno, measure.vars = traits, variable.name = "trait") #transformando o pheno em formato longo
head(pheno_long)

ggplot(pheno_long, aes(x = trait, y = value, color = trait)) + geom_boxplot(outlier.shape = NA)+
    geom_point(alpha = 0.2, position = "jitter", size = 1) + facet_grid(trait~LOCATION_NAME,
    scales = "free_y") + theme_bw() + theme(axis.text.x = element_text(angle=90))

library(ggh4x)
library(ggpmisc)

data_vis=ggplot(data = pheno_long, aes(x=trait, y= value, color=trait)) +
  geom_boxplot() +
  geom_point( position = "jitter", alpha = 0.5) +  facet_nested(trait~LOCATION_NAME+CROP_SEASON, scales = "free_y") +
  theme_bw();data_vis
```






```{r}
locations <-  pheno[,c("TRIAL_NAME","LOCATION_NAME","SITE_LAT","SITE_LONG")]
locations_unq <- locations |> distinct(SITE_LAT,.keep_all = TRUE)

locations_unq <- locations_unq[-21,]
smart_clean_coords <- function(x, decimal_places = 5) {
  # Replace commas with dots (fix decimals)
  x <- gsub(",", ".", x)
  
  # Remove thousand separators (dots between digits)
  x <- gsub("(?<=\\d)\\.(?=\\d)", "", x, perl = TRUE)
  
  # Convert to numeric (handles scientific notation like 0.00E+00)
  num <- as.numeric(x)
  
  # Function to scale down numbers that are too large
  scale_if_needed <- function(val) {
    if (is.na(val)) return(NA)
    
    # Latitude should be between -90 and 90
    if (abs(val) > 90) {
      # Keep dividing by 10 until in valid range
      while (abs(val) > 90) {
        val <- val / 10
      }
      return(val)
    } else {
      return(val)  # Already valid, return as-is
    }
  }
  
  # Apply scaling only where needed
  num <- sapply(num, scale_if_needed)
  
  # Round to specified decimal places
  round(num, decimal_places)
}
locations_unq$SITE_LAT <- smart_clean_coords(locations_unq$SITE_LAT)
locations_unq$SITE_LONG <- smart_clean_coords(locations_unq$SITE_LONG)

write.csv(locations_unq,file = "data/loc_unq.csv",row.names = F)
```

# Mapping the trials
```{r}
library(raster)
library(RColorBrewer)

#Escolha sua pasta de trabalho Control+Shift + H

#Importando o Brasil
Africa<- shapefile('data/Africa_Biomes_Dataset.shp')
Africa

#plot(brasil, axes=T, main= 'Brasil')


WGS84<-CRS("+proj=longlat +datum=WGS84 +no_defs ") #


# reg <- unique(brasil@data$REGIAO) #Obtendo o nome das regi?es
# (mreg <- match(brasil@data$REGIAO, reg))
# text(brasil, 'UF', cex=.6, halo=TRUE)


#Pacote Brewer Pal 
display.brewer.all(colorblindFriendly=T)

(paleta.colbrewer<-brewer.pal('Blues', n=5))
#paleta<- c("blue","red","green","magenta","yellow")

#plot(brasil, col=paleta[mreg], axes=T)
plot(Africa, col=paleta.colbrewer, axes=T)
text(brasil, 'UF', cex=.6, halo=TRUE)

pontos<- read.csv("data/loc_unq.csv", header=TRUE, sep=',') # se estiver usando outro OS
table(pontos$SITE_LONG)
head(pontos)
tail(pontos)
str(pontos)
# pontos[which(pontos$latitude < -23),]
# pontos[which(pontos$Instituicao=="CATI/Avare"),"regiao"]="SE"
# pontos[which(pontos$Instituicao=="DETEC/Itai"),"regiao"]="SE"
# pontos[which(pontos$cidade=="Campos Novos"),"regiao"]="SU"



# Apply to your data (5 decimal rounding)
pontos$SITE_LAT <- smart_clean_coords(pontos$SITE_LAT)
pontos$SITE_LONG <- smart_clean_coords(pontos$SITE_LONG)

# Check results
head(pontos)
tail(pontos)

coordinates(pontos)<- ~ SITE_LONG + SITE_LAT

WGS84<-CRS("+proj=longlat +datum=WGS84 +no_defs ") #
#plot(pontos,pch=20,cex=0.5, add=T)
crs(pontos)<-WGS84


# fv <- unique(pontos$regiao) 
# (m <- match(pontos$regiao, fv)) 
plot(pontos, bg=rainbow(4),pch=21, cex=0.75,add=T) 

#inserindo legenda
legend("bottomright",title ='Regions' ,legend = fv,cex = 1,x.intersp = 1,pch = 21, pt.cex = 2,pt.bg = rainbow(4))


#EXTRAS
#Plotando o norte
library(png)
library(grid)

norte<- readPNG('Norte.png')
grid.raster(norte, x=.7, y=0.8, width=.05)

#Escala
scalebar(d=5000,type='bar',divs=4)

agro<- raster('data/agro.tif') # Importando o Raster de elevação
hwsd <- raster('data/hwsd.tif')

plot(hwsd, main="Elevation") # Plotando o Raster de elevação
plot(agro, main="agro",col=rainbow(30))

palete.spec<- brewer.pal("Spectral",n=11)
plot(agro, main="Elevation", col= rev(palete.spec)) 

plot(africa, add=TRUE) #Adicionando o mapa do brasil


#Selecionando os estados

#brasil[LINHA , COLUNA]
# brasil
# 
# uf<- brasil[brasil@data$UF %in% c('MT','GO','RS','PE','BA','ES','DF',
#             'RJ','AL','SE','MG','MS','SP','PR','SC'),] #Mesma indexacao de data.frame
# uf
# 
# plot(uf, col='orange')
# plot(uf, col=c("red","blue"))
# text(uf,'UF', cex=1, halo=T)

#Cortando um raster para o estado

#for( i in 1:5){
agro_crop<- crop(agro,pontos)
plot(agro_crop)
hwsd_crop<- crop(hwsd,pontos)
plot(hwsd_crop)

# elev_uf<-mask(elev_uf,uf)
# plot(elev_uf, col=rev(palete.spec))

writeRaster(elev_uf,filename= "./UF_elev", overwrite=TRUE)
writeRaster(elev_uf,filename= "./UF_elev", overwrite=TRUE)
#raster::writeRaster(elev_uf,filename = paste0('Raster/r',i))
#}


plot(pontos, bg=rainbow(4)[m],pch=21, cex=0.75,axes=T,add=T) 


#Juntando tudo
#plot(uf, col='orange',axes=TRUE)
plot(elev_uf, col=rev(palete.spec), legend=F)
plot(brasil, add=T)
plot(pontos, bg=rainbow(4)[m],pch=21, cex=0.75,axes=T,add=T) 
legend("bottomright",pch=21,pt.cex=2, pt.bg=rainbow(4),legend = fv,cex = 1,y.intersp = 1.1, title = 'Brazillian regions',inset =0.01 )
text(brasil, 'UF', cex=.6, halo=TRUE)

```


## ggplot2
```{r,eval=FALSE}
#Com ggplot2

# elevação

library(raster)
library(RColorBrewer)
library(ggplot2)
library(sf)

africa.sf<-st_read("data/afr_g2014_2013_0.shp")
# uf.sf<- brasil.sf[brasil.sf$UF %in% c('MT','GO','RS','PE','BA','ES','DF',
#             'RJ','AL','SE','MG','MS','SP','PR','SC'), ]

library(stringi)
africa.sf$ADM0_NAME <- stri_replace_all_fixed(
  africa.sf$ADM0_NAME,
   "C\xf4te d'Ivoire",  # Literal special character
  "Cote d'Ivoire"
)
# elevação
#Selecionando os estados

africa<- shapefile('data/afr_g2014_2013_0.shp')
africa
#brasil[LINHA , COLUNA]


uf<- africa[africa@data$ADM0_NAME%in% c("Zimbabwe"),] #Mesma indexacao de data.frame



# # uf
# 
 plot(uf, col='orange')
# plot(uf, col=c("red","blue"))
# text(uf,'UF', cex=1, halo=T)

#Cortando um raster para o estado
agro<- raster('data/agro.tif') # Importando o Raster de elevação
hwsd <- raster('data/hwsd.tif')

plot(agro)
plot(hwsd)
#for( i in 1:5){
hwsd_africa<- crop(africa,uf)
plot(hwsd_africa)

hwsd_africa<-mask(africa,uf)
plot(hwsd_africa)
palete.spec<- brewer.pal("Spectral",n=11)
#plot(elev_uf, col=rev(palete.spec))

writeRaster(hwsd,filename= "data/hwsd_map", overwrite=TRUE)
writeRaster(agro,filename= "data/agro_map", overwrite=TRUE)
WGS84<-CRS("+proj=longlat +datum=WGS84 +no_defs ") #
#pontos
pontos<- read.csv("data/loc_unq.csv", header=TRUE, sep=',') # se estiver usando outro OS
str(pontos)
# pontos[which(pontos$latitude < -23),]
# pontos[which(pontos$Instituicao=="CATI/Avare"),"regiao"]="SE"
# pontos[which(pontos$Instituicao=="DETEC/Itai"),"regiao"]="SE"
# pontos[which(pontos$cidade=="Campos Novos"),"regiao"]="SU"

# Apply to your data (5 decimal rounding)
# pontos$SITE_LAT <- smart_clean_coords(pontos$SITE_LAT)
# pontos$SITE_LONG <- smart_clean_coords(pontos$SITE_LONG)

# Check results
head(pontos)
tail(pontos)

coordinates(pontos)<- ~ SITE_LONG + SITE_LAT
str(pontos)

#plot(pontos,pch=20,cex=0.5, add=T)
crs(pontos)<-WGS84


# fv <- unique(pontos$regiao) 
# (m <- match(pontos$regiao, fv)) 

#########
### Plotando o mapa do Brasil 
#

ggplot() + 
  geom_sf(data = africa.sf)
  #scale_fill_discrete(type = c("blue",'yellow','red','magenta','green'))

#########
### Plotando o mapa do Brasil  e da regiao cortada
#

ggplot() + 
  geom_sf(data = africa.sf, fill="white")


pontos.sf
pontos
#########
### Aproximando para a regi?o cortada
#


pontos@coords.nrs
(bbox_uf<-st_bbox(uf)) #Pegando os limites da regi?o cortada
ggplot() + 
  geom_sf(data = africa.sf, fill="white") +
  #Código para definir os limites do mapa cood_sf()
    coord_sf(xlim =c(bbox_uf["xmin"],bbox_uf["xmax"]) ,ylim=c(bbox_uf["ymin"],bbox_uf["ymax"]))+
   theme_light()
  # 
  

#########
### Adicionando os rótulos de estado
#

ggplot() + 
  geom_sf(data = africa.sf, fill="grey") +
  theme_light()
 
#########
### Plotando pontos
#

(pontos.sf<-data.frame(pontos)) #retornando nossos pontos para tabela

# pontos.sf <-pontos.sf[-21,] 
pontos.sf
# ggplot() + 
#   geom_sf(data = brasil.sf, fill="grey") +
#   geom_sf(data = uf.sf,fill="white")+
#   coord_sf(xlim =c(bbox_uf["xmin"],bbox_uf["xmax"]) ,ylim=c(bbox_uf["ymin"],bbox_uf["ymax"])) +
#   geom_sf_label(data = brasil.sf, aes(label = UF),fill=NA,label.size = NA)+
#   #Código para plotagem de pontos (geom_point)
#   geom_point(data = pontos.sf, mapping = aes(x = longitude, y = latitude, color=regiao),size =1) + 
#   scale_color_manual(breaks = c("NE", "CO", "SE","SU"),values=rainbow(4))+
#   theme_light()+ guides(color = guide_legend(title = "Region"))+labs(y="Latitude",x="Longitude")


problems <- africa.sf$ADM0_NAME[!stringi::stri_enc_isascii(africa.sf$ADM0_NAME)]
if(length(problems) > 0) print(problems)

#########
### com rasters
#
# install.packages("tmap")
library(stars)

elev_uf.stars<- read_stars("data/hwsd_map.grd",RasterIO = list(nBufXSize = 600, nBufYSize = 600))
# elev_uf.stars<- read_stars("data/agro_map.grd",RasterIO = list(nBufXSize = 600, nBufYSize = 600))
names(elev_uf.stars)<-"Elevation"
pontos

st_bbox(africa)
(bbox_uf<-st_bbox(uf))
(pontos.sf<-data.frame(pontos)) 



mapa=ggplot() + 
  geom_sf(data = africa.sf, fill="white") +
  #Código para plotagem do rater geom_stars()+
  coord_sf(xlim =c(bbox_uf["xmin"],bbox_uf["xmax"]) ,ylim=c(bbox_uf["ymin"],bbox_uf["ymax"]),crs = "WGS84") +theme_light()+
  geom_point(data = pontos.sf, mapping = aes(x = SITE_LONG, y = SITE_LAT),size =1,show.legend = T)+
    geom_sf_label(data = africa.sf, aes(label = ADM0_NAME),fill=NA,label.size = NA)+theme(legend.position = "none",axis.text.x =element_text(angle = 90))+labs(x="",y="");mapa


africamap=ggplot() + 
  geom_sf(data = africa.sf, fill="white")+
    scale_fill_distiller(palette = "RdYlGn",na.value = NA,name = "Altitude")+
  geom_stars(data = elev_uf.stars,alpha =0.8)+
  coord_sf(xlim =c(-25.35875,63.50265) ,ylim=c(-46.9813,  37.56095),crs = "WGS84")+labs(y="Latitude",x="Longitude")+
  theme_light()+theme(legend.position = "none")+
  geom_rect(data=africa.sf,
    xmin = 25.23703,
    ymin =-22.41774,
    xmax = 33.05630,
    ymax =-15.60883 ,
    fill = NA, 
    colour = "black",
    linewidth = 0.8
  );africamap


library(cowplot)

map=africamap |> ggdraw()+
  draw_plot(
    {
      mapa + 
        coord_sf(
          xlim = c(25.23703, 33.05630),
          ylim = c(-22.41774, -15.60883),
          expand = FALSE) +
        theme(legend.position = "none")
      },
    # The distance along a (0,1) x-axis to draw the left edge of the plot
    x = 0.13, 
    # The distance along a (0,1) y-axis to draw the bottom edge of the plot
    y = 0.05,
    # The width and height of the plot expressed as proportion of the entire ggdraw object
    width = 0.40, 
    height = 0.40);
map









ggsave(plot=map,device = "pdf",filename ="Plots/soybean_zimbabwe.pdf",dpi = "retina",height = 6,width = 8)
ggsave(plot=map,device = "png",filename ="Plots/soybean_zimbabwe.png",dpi = "retina",height = 6,width = 8)

```

# Single-environment analyses

These analyses had a purpose: to test the significance of the genetic variance in each environment. We used the following model:

$$
 \mathbf{y} = \mathbf{X_1b} + \mathbf{X_2che} + \mathbf{Z_1g} + \mathbf{\epsilon} 
$$

where $\mathbf{y}$ is the vector of phenotypic observations, $\mathbf{b}$ is the vector of fixed effects of replication, $\mathbf{che}$ is the vector of fixed effects of checks, $\mathbf{g}$ is the vector of random effects of genotypes  and $\mathbf{\epsilon}$ the vector of errors. $\mathbf{X_1}$, $\mathbf{X_2}$, $\mathbf{Z_1}$ $ are incidence matrix of $\mathbf{b}$, $\mathbf{g}$  effects respectively.

At each environment, we filtered out trials that do not have significance for genotype effect. The Likelihood Ratio Test ($LRT$) was computed as follows:

$$
LRT= −2 \times (Log𝐿 - Log L_𝑅)
$$

where $L$ is the maximum point of residual likelihood function of the complete model and $L_R$ is the same for the reduced model, that is, without the effect to be tested. The LRT value was compared with a tabulated value based on the chi-square table, with one degree of freedom and 0.95 probability.

We calculated the broad-sense heritability ($h^2$):

$$
h^2 = \sigma^2g    / \sigma^2g + \sigma^2e
$$

where $\sigma^2g$ is the is the additive genetic variance,$\sigma^2g$ is non additive genetic variance and $\sigma_e^2$ is the error variance.

## STMT

```{r first step,eval=FALSE, echo=TRUE}
source("https://raw.githubusercontent.com/saulo-chaves/May_b_useful/main/update_asreml.R")
#source("GEBV/codes/outliers.R")

library(asreml)

data.list = split(pheno, f = pheno$trial)

traits
stmt = list()
j="Banket_2018" 
#j="2021NAC_ST01"
i="Plant lodging"
unique(pheno$trial)


for (j in names(data.list)) {
  
   x = data.list[[j]]
  cat("====> Environment:", j, fill = TRUE)
  


 #   A.env = full2sparse(K = amat[which(rownames(amat) %in%
 #                                              c(x$geno)),
 #                                     which(colnames(amat) %in%
 #                                            c(x$geno))])
 # attr(A.env, "INVERSE") = FALSE
 
#  H.env.van = full2sparse(K = H.inv.van[which(rownames(H.inv.van) %in%
#                                               c(x$geno)),
#                                      which(colnames(H.inv.van) %in%
#                                             c(x$geno))])
#  
# attr(H.env.van, "INVERSE") = TRUE


  # x = x |> mutate(nad=geno)  |> mutate_at(vars(env, rep, geno, row, col, check,nad), as.factor)
  
  summar.mat = matrix(nrow = length(traits), ncol = 4, 
                      dimnames = list(traits, 
                                      c("geno","error","H2", "CV")))

  vc.env = list()
  Blue.env = list()
  
  for (i in traits) {
    
    cat("====> Trait:", i, fill = TRUE)
    
    if(all(is.na(x[,i]))) next
    
    model.r = asreml(fixed = x[,i] ~ rep, 
                   random = ~ geno,maxit = 50,
                   data = x, na.action = na.method(x = "include", y = "include"))
    

    
   # colocar nugget and spline
    model.f = asreml(fixed = x[,i] ~ rep+geno, 
                    maxit = 50,
                   residual = ~units,
                   data = x, na.action = na.method(x = "include", y = "include"))
 
    if(!model.r$converge) next
    
    if(any(na.exclude(model.r$vparameters.pc > 1))) model.r = update.asreml(model.r)
    # varcomp = summary(modelhvan)$varcomp
    varcomp = summary(model.r)$varcomp
      varcomp.df = data.frame(
      effect = c('genotypic', "residual"),
      component = c(round(
        varcomp[grep('geno', rownames(varcomp)),1],4),
        round(varcomp[grep('!R', rownames(varcomp)),1],4))
    )
     # if(!modelad$converge) next
     # 
     # if(any(na.exclude(modelad$vparameters.pc > 1))) modelad = up.mod(modelad)
     # 
    if(!model.f$converge) next
    
    if(any(na.exclude(model.f$vparameters.pc > 1))) model = up.mod(model.f)
      
    lrtest = as.data.frame(lrt(asreml(fixed = x[,i] ~ rep, 
                        residual = ~units, maxit = 50,
                        data = x),model.r))
    lrtest=up.mod(lrtest)
      
    lrtest$effect = c("genotypic")
    
    varcomp= full_join(varcomp.df, lrtest[,-1], by = "effect")
    varcomp$signi = ifelse(varcomp$`Pr(Chisq)` <= 0.06, "*", "ns")
    varcomp$signi2 = ifelse(varcomp$`Pr(Chisq)` <= 0.06, 
                            paste(round(varcomp$component, 4), "*"), 
                            paste(round(varcomp$component, 4), "ns"))
    rownames(varcomp)=varcomp$effect
    
    pred = predict(model.r, classify = "geno", sed = TRUE)
    pred$sed= pred$sed[-which(pred$pvals$status=="Aliased"),-which(pred$pvals$status=="Aliased")]
    MVdelta = mean((pred$sed^2)[upper.tri(pred$sed^2,diag = F)])
    H2 = 1-(MVdelta/(2*varcomp[varcomp$effect == "genotypic","component"]))
    h2 = vpredict(model.r, h2~(V1)/(V1+V2))



    #h2ad = vpredict(modelad, h2~(V3)/(V3+V4))
 
    CV = sqrt(varcomp[grep("r", rownames(varcomp)),"component"])/
      mean(x[,i], na.rm = TRUE)*100
    
    summar.mat[i,] = c(varcomp$signi2[1], round(varcomp$component[2],4), 
                       round(h2$Estimate,4), round(CV,4))
    
    
    pred=predict(model.f, classify = "geno", vcov = T)


    pred$vcov= pred$vcov[-which(pred$pvals$status=="Aliased"),-which(pred$pvals$status=="Aliased")]
    pred$pvals= pred$pvals[-which(pred$pvals$status=="Aliased"),]

    pred$pvals$w=diag(solve(pred$vcov))
    pred=pred$pvals
    
  vc.env[[i]] = varcomp.df

  Blue.env[[i]] = pred
  }
  stmt[[j]] = list(Blues = Blue.env,summary = summar.mat,vc=vc.env)
  
  rm(summar.mat,GEBV.env, varcomp, lrtest, model, 
     h2, CV, x, H.inv.env)
}

saveRDS(stmt, file = "Saves/stmt.rds")
#rm(data.list)

#hist(data[which(data$env=="UGN2021ABI_ST02"),"rytha"])
#table(data[which(data$env=="UGN2021ABI_ST02"),"rytha"],useNA = "ifany")

#table(data[which(data$env=="UGN2020NAS_ST04"),"vir2"], useNA = "ifany")

```
### Results
```{r}
stmt=readRDS("Saves/stmt.rds")

stvc=lapply(stmt, function(x)as.data.frame(cbind(x$summary)))

#stvc.df=lapply(stmti, function(x)as.data.frame(cbind(x$vc)))



#stmt[[1]][["Blues"]][[1]][,-c(3,4)]


stvc=do.call(rbind,stvc)
stvc$trait=rownames(stvc)
stvc$env=sub("\\..*","",stvc$trait)
stvc$trait = sub(".*\\.","",stvc$trait)
stvc$geno_sig=sub("^\\S+\\s+","",stvc$geno)
stvc$geno=sub("\\s.*","",stvc$geno)
stvc$geno= as.numeric(stvc$geno)
stvc$H2=as.numeric(stvc$H2)
stvc$error=as.numeric(stvc$error)
stvc$CV=as.numeric(stvc$CV)
library(tidyverse)
head(stvc)
stvcl= stvc |> select(geno, error, trait, env, geno_sig) |> pivot_longer(geno:error) 
  

unique(stvc$env)
# cols.label=c("A20NAC", "A20NAS", "B20NAS", "A21ABI" ,"B21ABI","A21NAC", "B21NAC", "A21NAS", "B21NAS")
# names(cols.label)=c(unique(stvc$env)) 

stvcl|>  ggplot(aes(x=name,y=value))+ geom_col(data= subset(stvcl, !grepl("ns", stvcl$geno_sig) & grepl("geno",stvcl$name)),aes(fill="*"), color="black", fill="#a8ddb5") + geom_col(data= subset(stvcl,  grepl("error",stvcl$name)),color="black",fill="lightblue") + facet_grid(trait~env,scales = "free_y") + geom_text(data=subset(stvcl, !grepl("ns", stvcl$geno_sig) & grepl("geno",stvcl$name)),aes(label = geno_sig), size=6 ) + theme(legend.position = "top", )  + geom_col(data= subset(stvcl, grepl("ns", stvcl$geno_sig) & grepl("geno",stvcl$name)),aes(fill="ns"), fill="gray", color="black")  + theme(legend.position = "top", ) + labs(x="Effects", y="Genotypic variance component") +   geom_text(data=subset(stvcl, grepl("ns", stvcl$geno_sig) & grepl("geno",stvcl$name)),aes(label = "ns", vjust = -1.0), size=3 ) 

View(stvcl)

stvc_bef <- stvc |>  select(H2, CV, trait, env, geno_sig) |> pivot_longer(H2:CV) |>  
  ggplot(aes(x=env,y=value))+geom_col(aes(fill = trait), position = position_dodge())+ facet_grid(name~trait, scales = "free_y") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) + labs(x= "Environments", y="Parameters")

stvc$CV[stvc$CV == 948.7212] <- NA
stvc$CV[stvc$CV == 130.8657] <- NA
stvc$CV[stvc$CV == 129.7948] <- NA

stvc_aft <- stvc |>  select(H2, CV, trait, env, geno_sig) |> pivot_longer(H2:CV) |>  
  ggplot(aes(x=env,y=value))+geom_col(aes(fill = trait), position = position_dodge())+ facet_grid(name~trait, scales = "free_y") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) + labs(x= "Environments", y="Parameters")

   z = data.list[["Stapleford_2018"]]
   z = data.list[["Harare_2021"]]
   table(z$`Plant lodging`)
   z = data.list[["Harare_2020"]]
   table(z$`Plant lodging`)
    z = data.list[["Harare_2022"]]
   table(z$`Plant lodging`)
   
   
   stvc |> mutate(CV = case_when(CV == 948.7212~NA))
```

## preparing MT data
```{r mtdata, results='hide', eval=FALSE}
traits 
stmt = readRDS(file = "Saves/stmt.rds")

str(stmt$Banket_2019)
stmt$Banket_2019$Blues$`Plant lodging`
i
stmt$Banket_2018$Blues$`Plant lodging`$
mtdata = list()
diagnostic

#i="Banket_2018"
#j="Plant lodging"
stmt$Banket_2018$Blues
for (i in names(stmt)) {
  mtdata[[i]] = stmt[[i]][["Blues"]][["Plant lodging"]][,-c(3,4)]
  if(is.null(mtdata[[i]])){
    mtdata[[i]] = stmt[[i]][["Blues"]][[j]][,-c(3,4)]
    colnames(mtdata[[i]]) = c("geno", j, paste('w',j,sep = "_"))
  mtdata[[i]]$env = i
  mtdata[[i]] = mtdata[[i]][,c("geno","env",j, paste('w',j,sep = "_"))]
  }else{
  colnames(mtdata[[i]]) = c("geno", "Plant lodging", "w_Plant lodging")
  mtdata[[i]]$env = i
  mtdata[[i]] = mtdata[[i]][,c("geno","env","Plant lodging", "w_Plant lodging")]
    }
  for (j in traits.sel[-1]) {
    if(print(all(mtdata$geno == stmt[[i]][["Blues"]][[j]][,1]))) cat("The names match", fill = TRUE)
    
    mtdata[[i]][,c(j, paste('w',j,sep = "_"))] = stmt[[i]][["Blues"]][[j]][,c(2,5)]

  }
}

mtdata[[3]]$vir2=NA
mtdata[[3]]$w_vir2=NA
mtdata[[5]]$rytha=NA
mtdata[[5]]$w_rytha=NA
mtdata[[5]]=mtdata[[5]][,c("geno", "env" ,"Plant lodging", "w_Plant lodging" ,"Plant Height", "w_Plant Height" , "Grain Yield","w_Grain Yield")]

met_data = do.call(rbind, mtdata); rownames(met_data) = NULL

rm(mtdata)
```

```{r, eval=FALSE}

stmt = readRDS(file = "Saves/stmt.rds")

diagnostic = do.call(rbind, lapply(stmt, function(x){
  aa = as.data.frame(x$summary) |> rownames_to_column("trait")
  aa
})) |> rownames_to_column("env") |> 
  mutate_at("env", str_replace, "\\..*", "") |> 
  dplyr::select(env, trait, geno) |> 
  separate(geno, into = c("genovar", "sig"), sep = " ")

# diagnostic
# dim(pheno)
# dim(pheno_filt)
# unique(pheno_filt$trial)
# levels(pheno_filt$trial)


# inpheno# install.packages("ProbBreed")
library(ProbBreed)
# head(pheno_filt)

i=traits[1]

# Getting the path of your current open file
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
print( getwd() )

for (i in traits) {
  message(paste("Analysis to", i))
  # print(paste("Analysis to", i)) 
  # cat("====> Trait:", i, fill = TRUE)
  pheno_filt <- droplevels(pheno[
    which(pheno$trial %in% diagnostic[which(diagnostic$trait == i &
                                               diagnostic$sig == "*"), "env"]), ])
  

   
  dim(pheno_filt)
  mod = bayes_met(data = pheno_filt,
                gen = "geno",
                loc = "LOCATION_NAME",
                repl = c("rep"),
                trait = paste(i),
                reg = NULL,
                year = NULL,
                res.het = TRUE,
                iter = 4000, cores = 4, chain = 4)
  
#save(mod,file = paste0("Saves/",i,".rda"))  
saveRDS(mod,file = paste0("Saves/",i,".rds"))
rm(mod)
}

rm(list=ls())

mod_PH=readRDS("Saves/Plant Height.rds")
mod_PL=readRDS("Saves/Plant lodging.rds")
mod_GY=readRDS("Saves/Grain Yield.rds")



outs_PH = extr_outs(model = mod_PH,
                 probs = c(0.05, 0.95),
                 verbose = TRUE)

outs_PL = extr_outs(model = mod_PL,
                 probs = c(0.05, 0.95),
                 verbose = TRUE)
outs_GY = extr_outs(model = mod_GY,
                 probs = c(0.05, 0.95),
                 verbose = TRUE)

saveRDS(outs_PH,file = "Saves/outs_PH.rds")
saveRDS(outs_PL,file = "Saves/outs_PL.rds")
saveRDS(outs_GY,file = "Saves/outs_GY.rds")

setwd("./Saves/PH")
res_PH = prob_sup(extr = outs_PH,
                   int = .2,
                   increase = FALSE,
                   save.df = FALSE,
                   verbose = FALSE)
setwd("../PL")

res_PL = prob_sup(extr = outs_PL,
                   int = .2,
                   increase = FALSE,
                   save.df = FALSE,
                   verbose = TRUE)
setwd("../GY")

res_GY = prob_sup(extr = outs_GY,
                   int = .2,
                   increase = TRUE,
                   save.df = TRUE,
                   verbose = FALSE)

setwd("../")

save(res_PL,file = "res_PL.rda")
save(res_PH,file = "res_PH.rda")
save(res_GY,file = "res_GY.rda")

```



# outs$ppheck
```{r}
library(ProbBreed)
library(tidyverse)
outs_PH=readRDS("Saves/outs_PH.rds")
outs_PL=readRDS("Saves/outs_PL.rds")
outs_GY=readRDS("Saves/outs_GY.rds")


outs_PH$ppcheck
outs_PL$ppcheck
outs_GY$ppcheck

plot(outs_PH, category = "density")
# plot(outs_PH, category = "histogram")

plot(outs_PL, category = "density")
# plot(outs_PL, category = "histogram")

plot(outs_GY, category = "density")
# plot(outs_GY, category = "histogram")
# Density: Empirical vs Sampled phenotypic values
PL=plot(outs_PL)+ggtitle("")
PH=plot(outs_PH)+ggtitle("")
GY=plot(outs_GY)+ggtitle("")
library(ggpubr)
plot.density = ggarrange(PH,PL,GY,labels = c("A","B","C"),nrow = 3, common.legend = T)


ggsave(plot=plot.density,device = "pdf",filename ="Plots/plots_density.pdf",dpi = "retina",height = 6,width = 8)
ggsave(plot=plot.density,device = "png",filename ="Plots/plots_density.png",dpi = "retina",height = 6,width = 8)

```

```{r}

load("Saves/res_PL.rda")
load("Saves/res_PH.rda")
load("Saves/res_GY.rda")

head(selected)
intersect(selected$gen,bp.sel)
head(gen_sel)
gen_sel_l=gen_sel |> pivot_longer(prob_gy:prob_cga:prob_pa) 
head(gen_sel_l)
 gen_sel_l |> 
  ggplot(aes(x = gen, y = value, 
             fill = ifelse(gen %in% selected$gen, "selected", "non-selected"))) +
  geom_bar(stat = "identity", position = "stack") +
  geom_bar(data = subset(gen_sel_l, gen %in% selected$gen), 
           stat = "identity", position = "stack", fill = "lightblue") +
  facet_wrap(~name, scales = "free_x") + 
  geom_text(data = subset(gen_sel_l, gen %in% selected$gen), 
            aes(label = value), colour = "black", vjust = 1.0) +
  theme(axis.text.x = element_text(size = 5, angle = 90, hjust = 1, vjust = 0.5),
    panel.background = element_blank(),
    legend.position = "top", 
    strip.text = element_text(size = 14)
  ) +
  scale_fill_manual(values = c("selected" = "lightblue",
                               "non-selected" = "gray")) +  
  labs(x = "Carioca common bean families", y = "Ranking", 
       fill = expression(bold(Pr(g %in% Omega)))) +
  theme(
    axis.text.x = element_text(size = 5, angle = 90, hjust = 1, vjust = 0.5),
    panel.background = element_blank(),
    legend.position = "top", 
    strip.text = element_text(size = 14)
  ) + scale_y_reverse() 

PH$data$

plot(res_PL)+ggtitle("PL")+ geom_text(
    data = subset(df, ID %in% gen.sel),  # Filter points to label
    aes(label = ID),                     # Column containing labels
    colour = "black",
    vjust = 1.0
  )
PH=plot(res_PH)+ggtitle("PH")
GY=plot(res_GY)+ggtitle("GY")
library(ggpubr)
plot.perfo = ggarrange(PH,PL,GY,labels = c("A","B","C"),nrow = 3, common.legend = T)


ggsave(plot=plot.density,device = "pdf",filename ="Plots/plots_density.pdf",dpi = "retina",height = 6,width = 8)
ggsave(plot=plot.density,device = "png",filename ="Plots/plots_density.png",dpi = "retina",height = 6,width = 8)

```


# BPSI
```{r}

load("Saves/res_PL.rda")
load("Saves/res_PH.rda")
load("Saves/res_GY.rda")

source("data/bpsi_fun.R")

models= vector("list",length(3))


models[[1]] = res_GY;
models[[2]] = res_PH
models[[3]] = res_PL
models[[3]]$across$perfo <- models[[3]]$across$perfo[-which(models[[3]]$across$perfo$ID=="G_55"),]
names(models) <- c("GY","PH","PL")
BPSI_soy=BPSI(modlist = models,increase =c(TRUE,FALSE,FALSE),omega = c(2,1,1),int = 0.1,save.df = T,verbose = T )

df=print(BPSI_soy)

gen.sel = df[which(df$sel=="selected"),"gen"]

```

# BPSI plot

Highlighting the ranking of probability of superior performance per trait of the selected families.

```{r bpsi plot ranks}
#| label: fig-ranks
#| fig-cap: "Ranking of probability of superior performance per trait"
library(knitr)

plot(BPSI_soy,category = "Ranks")
```


```{r bpsi plot}
#| label: fig-bpsi
#| fig-cap: "Selected families based on BPSI rank"
plot(BPSI_soy,category = "BPSI")
library(kableExtra)

BPSI_soy$BPSI |> kable(format = "latex", booktabs = TRUE, longtable=TRUE)

BPSI_soy$BPSI |> kable(format = "html", booktabs = TRUE, longtable=TRUE)

```

# plots
```{r}

load("Saves/res_PL.rda")
load("Saves/res_PH.rda")
load("Saves/res_GY.rda")

res_PL$across$stabi


models=list(mod,mod2,mod3)
#models <- vector("list",length(mods))
names(models) <- c("GY","LOD","CB")
#str(models)
outs= vector("list",length(models))
names(outs) <- c("GY","LOD","CB")



```


```{r}

```


\pagebreak
