---
title: "BPSI pipeline"
author: "José Tiago"
date: "2024-10-26"
output:
  pdf_document:
    toc: true
    toc_depth: 3
  word_document:
    toc: true
    toc_depth: '3'
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
bibliography: references.bib
editor_options:
  chunk_output_type: console
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
---

```{r, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, tidy = TRUE, warning = FALSE, 
message = FALSE,   fig.width = 10, fig.height = 7)
#knitr::opts_knit$set(root.dir = "./")
```



\pagebreak



# Bayesian Probabilistic Selection Index (BPSI)


BPSI is a selection index that allows select genotypes through the probability of superior performance across environments @Dias2022.




$$
Pr(\hat{g}_i \in \Omega | y) = \frac{1}{S} \sum_{s=1}^SI(\hat{g}_i^{(S)} \in \Omega|y)
$$

 
 
where $\hat{g}_i$ is the posterior genotypic value, $\Omega$ is the superior group given the desired selection intensity, S represents the number of samples from the posterior distribution, and I $\left({\hat{g}}_i^{\left(s\right)}\ \in \Omega\middle|y\right)$ is an indicator variable for success (1) or failure (0). Selection intensity was set at 20%. 


We run the Bayesian model for each trait and after bind all them in a probabilistic rank of superior performance.



# Bayesian model


The Bayesian model was carried out in ProbBreed package @Chaves2023 .


``` {r, eval=FALSE,echo=TRUE}

library(ProbBreed)
library(tidyverse)
head(met_df)
met_df=met_df |> rename("GY"=7,"LOD"=8,"CB"=9) 
head(met_df)
mod = bayes_met(data = met_df, 
                gen = "id",
                loc = "env",
                repl = c("block"),
                trait = "GY", 
                reg = NULL,
                year = NULL, 
                res.het = T, 
                iter = 4000, cores = 4, chain = 4)
saveRDS(mod,file = "Simulation data/mod_T1S1.rds")

mod2 = bayes_met(data = met_df, 
                gen = "id",
                loc = "env",
                repl = c("block"),
                trait = "LOD", 
                reg = NULL,
                year = NULL, 
                res.het = T, 
                iter = 4000, cores = 4, chain = 4)
saveRDS(mod2,file = "Simulation data/mod_T2S1.rds")

mod3 = bayes_met(data = met_df, 
                gen = "id",
                loc = "env",
                repl = c("block"),
                trait = "CB", 
                reg = NULL,
                year = NULL, 
                res.het = T, 
                iter = 4000, cores = 4, chain = 4)
saveRDS(mod3,file = "Simulation data/mod_T3S1.rds")
#mod3=readRDS("Simulation data/mod_T3S1.rds")
#attr(mod3,"data")

```


# Probability of superior performance


The Probbreed package helped to estimate the probabilities for the BPSI ranks @Chaves2024 .

In this example we simulated 200 maize genotypes evaluated for grain yied, lodging and culm break in three environments. We used the packages alphasimR (@Gaynor2021) and fieldsimR (@Werner2024) to simulate genotype and phenotype effects respectively.


``` {r BPSI, eval=F, echo=TRUE}
rm(list=ls())
library(ProbBreed)
library(tidyverse)

models=list(mod,mod2,mod3)
#models <- vector("list",length(mods))
names(models) <- c("GY","LOD","CB")
#str(models)
outs= vector("list",length(models))
names(outs) <- c("GY","LOD","CB")

for (mods in 1:length(models)) {
  
  a= extr_outs(model = models[[mods]],
                 probs = c(0.05, 0.95),
                 verbose = TRUE)
  outs[[mods]]=a
  rm(a)
}
results= vector("list",length(outs))
names(results) <- c("GY","LOD","CB")

inc=c(TRUE,FALSE,FALSE)
for (mods in 1:length(outs)) {
  
  a = prob_sup(extr = outs[[mods]], 
                   int = .2,
                   increase = inc[[mods]], 
                   save.df = FALSE, 
                   verbose = TRUE)
  results[[mods]]=a
  rm(a)
}

saveRDS(results,file = "Data/resultsS1.rds")



```



# Applying BPSI to example dataset

```{r rank}
results=readRDS("Data/resultsS1.rds")

source("Data/bpsi_fun.R")

bpsi=BPSI(modlist=results, increase = c(TRUE,FALSE,FALSE),int = 0.1,omega=c(2,1,1),save.df = F,verbose = T)
```


# BPSI plot

Highlighting the ranking of probability of superior performance per trait of the selected families.

```{r bpsi plot ranks}
#| label: fig-ranks
#| fig-cap: "Ranking of probability of superior performance per trait"
library(knitr)

plot(bpsi,category = "Ranks")
```


```{r bpsi plot}
#| label: fig-bpsi
#| fig-cap: "Selected families based on BPSI rank"
plot(bpsi,category = "BPSI")
library(kableExtra)

bpsi$BPSI |> kable(format = "latex", booktabs = TRUE, longtable=TRUE)


```



\pagebreak

# References
